In this question you will fit various spline models to the fossil data (in "fossil.csv") of Chaud-huri and Marron (1999). These data consist of 106 measurements of ratios of strontium isotopes found in fossil shells and their ages.

```{r}
library(splines)
library(gam)
library(ISLR)
library(tidyverse)
fossil <- read_csv(file = "fossil.csv")
```


(a) Fit a polynomial function, using 5 fold cross-validation for choosing the degree of the polynomial. Report the selected model.

```{r}
library(rsample)
folds <- vfold_cv(fossil, 5)
folds %>% mutate(fits = splits %>% map(~lm(age ~ poly(strontium.ratio, 2), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds2 <- vfold_cv(fossil, 5)
folds2 %>% mutate(fits = splits %>% map(~lm(age ~ poly(strontium.ratio, 3), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))
folds3 <- vfold_cv(fossil, 5)
folds3 %>% mutate(fits = splits %>% map(~lm(age ~ poly(strontium.ratio, 4), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds4 <- vfold_cv(fossil, 5)
folds4 %>% mutate(fits = splits %>% map(~lm(age ~ poly(strontium.ratio, 5), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds5 <- vfold_cv(fossil, 5)
folds5 %>% mutate(fits = splits %>% map(~lm(age ~ poly(strontium.ratio, 6), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))
```
A fourth-degree polynomial yields the lowest MSE. 

(b) Fit a natural cubic spline, using 5 fold cross-validation for choosing the degrees of freedom. Report the degrees of freedom of the chosen model.

```{r}
library(rsample)
library(dplyr)
folds <- vfold_cv(fossil, 5)
folds %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 1), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds1 <- vfold_cv(fossil, 5)
folds1 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 2), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds2 <- vfold_cv(fossil, 5)
folds2 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 3), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds3 <- vfold_cv(fossil, 5)
folds3 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 4), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds4 <- vfold_cv(fossil, 5)
folds4 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 5), data = analysis(.)))) %>%         mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds5 <- vfold_cv(fossil, 5)
folds5 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 6), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds6 <- vfold_cv(fossil, 5)
folds6 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 7), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds7 <- vfold_cv(fossil, 5)
folds7 %>% mutate(fits = splits %>% map(~ lm(age ~ ns(strontium.ratio, df = 8), data = analysis(.)))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))
```
So 7 degrees of freedom yields the lowest MSE.

(c) Fit a smoothing spline, using 5 fold cross-validation for choosing the tuning parameter. Report the degrees of freedom of the chosen model.

```{r}
fit_smooth <- smooth.spline(fossil$strontium.ratio, fossil$age, cv = TRUE, df = 6)
```


(d) Fit a local linear regression, using 5 fold cross-validation for choosing the tuning parameter. Report the corresponding degrees of freedom of the chosen model.

```{r}
folds <- vfold_cv(fossil, 5)
folds %>% mutate(fits = splits %>% map(~loess(strontium.ratio ~ age, data = analysis(.), span = 1))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds1 <- vfold_cv(fossil, 5)
folds1 %>% mutate(fits = splits %>% map(~loess(strontium.ratio ~ age, data = analysis(.), span = 2))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds2 <- vfold_cv(fossil, 5)
folds2 %>% mutate(fits = splits %>% map(~loess(strontium.ratio ~ age, data = analysis(.), span = 3))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds3 <- vfold_cv(fossil, 5)
folds3 %>% mutate(fits = splits %>% map(~loess(strontium.ratio ~ age, data = analysis(.), span = 4))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))

folds4 <- vfold_cv(fossil, 5)
folds4 %>% mutate(fits = splits %>% map(~loess(strontium.ratio ~ age, data = analysis(.), span = 5))) %>%
        mutate(mses = map2_dbl(splits, fits, ~ modelr::mse(.y, assessment(.x)))) %>%
        summarize(sum(mses))
```
1 degree of freedom yielded the smallest MSE. 


(e) Make a scatterplot of the data, superimposed with the fitted curves by (a), (b), (c) and (d). Compare and discuss the fitted models.

```{r}
poly_fit <- lm(age ~ poly(strontium.ratio, 5), data = fossil)
nat_fit <- lm(age ~ ns(strontium.ratio, df = 3), data = fossil)
lin_fit <- loess(strontium.ratio ~ age, data = fossil, span = 1)
ggplot(fossil) + geom_point(aes(strontium.ratio, age))
par(new = TRUE)
plot(poly_fit, col = "blue")
par(new = TRUE)
plot(nat_fit, col = "orange")
par(new = TRUE)
plot(lin_fit, col = "red")
```


The best fit is the cubic spline 
